---
title: "Visualizing various types of data"
subtitle: "Lecture 3"
date: "May 20, 2024"
format: 
  revealjs:
    self-contained: true
---

# Warm up

```{r}
#| echo: false

library(countdown)
ggplot2::theme_set(ggplot2::theme_gray(base_size = 16))
```


## Questions ?? {.smaller}


# From last time

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

## Violin plots

::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-1
#| fig-show: hide
#| code-line-numbers: "|4|5|8"
#| echo: true

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g
    )
  ) +
  geom_violin()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-1
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::

## Multiple geoms

::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-2
#| fig-show: hide
#| code-line-numbers: "9"
#| echo: true

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g
    )
  ) +
  geom_violin() +
  geom_point()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-2
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::

## Multiple geoms

::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-3
#| fig-show: hide
#| code-line-numbers: "9"
#| echo: true

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g
    )
  ) +
  geom_violin() +
  geom_jitter()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-3
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::

## Multiple geoms + aesthetics

::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-4
#| fig-show: hide
#| code-line-numbers: "6"
#| echo: true

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g,
    color = species
    )
  ) +
  geom_violin() +
  geom_jitter()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-4
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::

## Multiple geoms + aesthetics

::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-5
#| fig-show: hide
#| echo: true
#| code-line-numbers: "11-13"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g,
    color = species
    )
  ) +
  geom_violin() +
  geom_jitter() +
  theme(
    legend.position = "none"
  )
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-5
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::

## Multiple geoms + aesthetics

::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-6
#| fig-show: hide
#| code-line-numbers: "14"
#| echo: true

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g,
    color = species
    )
  ) +
  geom_violin() +
  geom_jitter() +
  theme(
    legend.position = "none"
  ) +
  scale_color_colorblind()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-6
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::



```{r}
#| echo: false

library(countdown)
ggplot2::theme_set(ggplot2::theme_gray(base_size = 16))
```



## Questions from previous material {.smaller}

-   **Is there any code in the videos that is not in the readings?** Yes and no.
    There is no substantial functionality introduced in the videos that is not also in the readings, however the examples in the videos are different than the ones in the reading.

-   **What are all of the `geom`s we need to know?** You don't need to "memorize" or even "know" all o the geoms available in the ggplot2 package, but you can find a list of them on the [ggplot2 cheat sheet](https://ggplot2.tidyverse.org/#cheatsheet) or on the [reference page](https://ggplot2.tidyverse.org/reference/index.html#geoms).

-   **Could you please clarify what situations it would be appropriate to use each geom function?** Today's topic!
    And think about it as "what plot should I make for which type of variable".


# Let's return to AE-02

## `ae-02-bechdel-dataviz` {.smaller}


Go to the project navigator in RStudio (top right corner of your RStudio window) and open the project called `ae`.
If there are any uncommitted files, commit them so you can start with a clean slate.

## Recap of AE {.smaller}

-   Construct plots with `ggplot()`.
-   Layers of ggplots are separated by `+`s.
-   The formula is (almost) always as follows:

```{r}
#| eval: false
#| echo: true

ggplot(DATA, aes(x = X-VAR, y = Y-VAR, ...)) +
  geom_XXX()
```

-   Aesthetic attributes of a geometries (color, size, transparency, etc.) can be **map**ped to variables in the data or **set** by the user, e.g. `color = binary` vs. `color = "pink"`.
-   Use `facet_wrap()` when faceting (creating small multiples) by one variable and `facet_grid()` when faceting by two variables.
