---
title: "RECAP: Linear regression with a single predictor"
subtitle: "Lecture 15"
date: "June 6, 2024"
format: 
  revealjs:
    self-contained: true
    footer: "[🔗 sta199-summer24.github.io](https://sta199-s24.github.io/) &nbsp;·&nbsp; [❓ Ask on Ed](https://edstem.org/us/courses/59071/discussion/)"
execute: 
  echo: true
editor_options: 
  chunk_output_type: console
---


```{r}
#| echo: false
#| message: false

library(countdown)
library(tidyverse)
library(fivethirtyeight) # for the fandango dataset

ggplot2::theme_set(theme_gray(base_size = 16))
```




## Goals

-   Review linear regression with a single predictor

## Setup

```{r}
#| label: load-pkg
#| message: false

library(tidyverse)
library(tidymodels)
```





# Linear regression with a single predictor



```{r data-prep}
#| echo: false

movie_scores <- fandango |>
  rename(
    critics = rottentomatoes, 
    audience = rottentomatoes_user
  )
```

## Data overview

From the Fandango data set, we want to model `audience` movie ratings based on `critics` ratings.

```{r data-overview}
#| echo: true

movie_scores |>
  select(critics, audience)
```


## Regression model {#regression-model-1}

A **simple linear regression model** is used to model the relationship between a quantitative outcome ($Y$) and a single quantitative predictor ($X$): $$\Large{Y = \beta_0 + \beta_1 X +\epsilon}$$


-   $\beta_1$: True slope of the relationship between $X$ and $Y$
-   $\beta_0$: True intercept of the relationship between $X$ and $Y$
- $\epsilon$: Error term


## Fandango data: least squares line



```{r}
#| echo: false
#| message: false

m <- lm(audience ~ critics, data = movie_scores)
ggplot(data = movie_scores, 
       mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "#325b74", se = FALSE, linewidth = 1.5) +
  labs(x = "Critics Score" , 
    y = "Audience Score") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks.x = element_blank(), 
    axis.ticks.y = element_blank()
    )
```

Where the solid line minimizes the sum of the squared residuals.

## Linear regression model output

```{r}
movies_fit <- linear_reg() |>
  fit(audience ~ critics, data = movie_scores)

tidy(movies_fit)
```

# Slope and intercept


## Interpreting slope & intercept

$$\widehat{\text{audience}} = 32.3 + 0.519 \times \text{critics}$$

::: incremental
-   **Slope:** For every one point increase in the critics score, we expect the audience score to be higher by 0.519 points, on average.
-   **Intercept:** If the critics score is 0 points, we expect the audience score to be 32.3 points.
:::

## Is the intercept meaningful?

✅ The intercept is meaningful in context of the data if

-   the predictor can feasibly take values equal to or near zero or
-   the predictor has values near zero in the observed data

. . .

🛑 Otherwise, it might not be meaningful!


